Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-04-16 bis_open_loop_planning.tex
[MM] constraints formulation (first try)

The model described in Sec.~\ref{chap:vehicle_model} is embedded into a minimum lap time problem which is briefly outlined below.

>>>>

The proposed approach requires the presence of $H+1$ versions of the matrix $\bP$ at each grid point. One of these versions is initialized with $\bar{\bP}_0$, while the others represent the evolution of matrices that were initialized at previous grid points, from 1 to $H$ steps earlier. This ensures that at each step it is possible to find a covariance matrix evolved for $H$ steps and use it to robustify the constraint. 
To adapt the problem, it is necessary to modify equations (\ref{eq:DOCPdP}) and (\ref{eq:DOCPconstraints}). In this case,  (\ref{eq:DOCPdP}) collects $H$ collocation and continuity equations and an initialization equation, while (\ref{eq:DOCPconstraints}) takes as input the most propagated version of the covariance matrix at the k-th step, here named $\bP^H_k$. The new formulation is reported:
\begin{alignat}{3}
\bzero       = & \; \bPsiP_k(\bmu_k,\bxi_k, \bu_k, \bP_{k-1}^{j-1},\bP_k^j,\bSi_k^j,\bz_k),
& \quad & k = 1,\ldots, N; \; j=0,\ldots,H \label{eq:DOCPdPopenloop} \\
0            \geq & \; h_i(\bmu_k, \bu_k, \bz_k) + \be_i(\bmu_k, \bu_k, \bP_k^j, \bz_k),
& \quad & k = 1,\ldots, N;\; i \in \calI \label{eq:DOCPconstraintsopenloop}
\end{alignat}
This process is illustrated in Figure \ref{fig:DOCPgrid}, where the dashed rectangle represent the $k$-th step at which the constraints are being formulated. At each grid point, two versions of the covariance matrix are highlighted: $\bP^0_k$ (red node), which is the version to be initialized, and $\bP^H_k$ (green node), which is the most propagated version used to formulate the robust constraints in equation (\ref{eq:DOCPconstraints}).

\begin{figure}
	\centering
	\begin{tikzpicture}[%
		smallnode/.style={%
			circle, draw, minimum size=3mm, inner sep=0pt
		},
		r_smallnode/.style={%
			circle, draw, minimum size=3mm, inner sep=0pt, red
		},
		g_smallnode/.style={%
			circle, draw, minimum size=3mm, inner sep=0pt, mygreen
		},
		every path/.style={->, thick},
		x=1.5cm, y=1.5cm
		]

		\newcommand{\Plab}[2]{\bP^{#2}_{#1}}
		\definecolor{mygreen}{RGB}{0 150 0}

		% Nodes
		% First row
		\node[smallnode, label={[label distance=-2pt]above:\protect\(\Plab{k-2}{H-2}\)}] (Pk-2H-2) at (1,0) {};
		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-1}\)}] (Pk-1H-1) at (2,0) {};
		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H}\)}] (PkH) at (3,0) {};
		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{0}\)}] (Pk+10) at (4,0) {};
		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{1}\)}] (Pk+21) at (5,0) {};

		% Second row
		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-3}\)}] (Pk-2H-3) at (1,1) {};
		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H-2}\)}] (Pk-1H-2) at (2,1) {};
		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{H-1}\)}] (PkH-1) at (3,1) {};
		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{H}\)}] (Pk+1H) at (4,1) {};
		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{0}\)}] (Pk+20) at (5,1) {};

		% dots row 
		\node[label={center,rotate=90:\(\dots\)}] (dots2) at (1, 1.7) {};
		\node[label={center,rotate=90:\(\dots\)}] (dots3) at (2, 1.7) {};
		\node[label={center,rotate=90:\(\dots\)}] (dots4) at (3, 1.7) {};
		\node[label={center,rotate=90:\(\dots\)}] (dots5) at (4, 1.7) {};
		\node[label={center,rotate=90:\(\dots\)}] (dots6) at (5, 1.7) {};

		% Third row 
		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H}\)}] (Pk-2H) at (1,2) {};
		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{0}\)}] (Pk-10) at (2,2) {};
		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k}{1}\)}] (Pk1) at (3,2) {};
		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{2}\)}] (Pk+12) at (4,2) {};
		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{3}\)}] (Pk+23) at (5,2) {};

		% Fourth row
		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k-2}{H-1}\)}] (Pk-2H-1) at (1,3) {};
		\node[g_smallnode, label={[label distance=-2pt]above:\(\Plab{k-1}{H}\)}] (Pk-1H) at (2,3) {};
		\node[r_smallnode, label={[label distance=-2pt]above:\(\Plab{k}{0}\)}] (Pk0) at (3,3) {};
		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+1}{1}\)}] (Pk+11) at (4,3) {};
		\node[smallnode, label={[label distance=-2pt]above:\(\Plab{k+2}{2}\)}] (Pk+22) at (5,3) {};

		% Curved arrows
		\draw[bend left=20] (Pk-2H-2) to (Pk-1H-1);
		\draw[bend left=20] (Pk-1H-1) to (PkH);
		\draw[bend left=20] (Pk+10) to (Pk+21);

		\draw[bend left=20] (Pk-2H-3) to (Pk-1H-2);
		\draw[bend left=20] (Pk-1H-2) to (PkH-1);
		\draw[bend left=20] (PkH-1) to (Pk+1H);

		\draw[bend left=20] (Pk-10) to (Pk1);
		\draw[bend left=20] (Pk1) to (Pk+12);
		\draw[bend left=20] (Pk+12) to (Pk+23);

		\draw[bend left=20] (Pk-10) to (Pk1);
		\draw[bend left=20] (Pk1) to (Pk+12);
		\draw[bend left=20] (Pk+12) to (Pk+23);

		\draw[bend left=20] (Pk-2H-1) to (Pk-1H);
		\draw[bend left=20] (Pk0) to (Pk+11);
		\draw[bend left=20] (Pk+11) to (Pk+22);

		% Dashed rectangle
		\draw[dashed, thick, rounded corners] (2.6,-.25) rectangle (3.4,3.5);

	\end{tikzpicture}
	\caption{}
	\label{fig:DOCPgrid}
\end{figure}
